\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-03-05 reviews/response.Rmd

We have removed the comma and elsewhere, and adjusted the language based on the comments.

> p2, l9: "grammar of graphics" - provide reference immediately here and not only later in the article and explain what the grammar of graphics is; also "piped into" may not be understood by outsiders

Done.

> Fig.1 (& related figs): I suppose you use a "3-class Dark2" color scheme from Rcolorbrewer? This does not work well when printed in gray scale: The colors become almost indistinguishable. ...

@@

We have changed the colour to red and blue picked from a "4-class RdYlBu" colour scheme, which is different from the colours used for the 3 stations.

> Fig.2: Instead of "Jan 2016" ... "Jan 2017", be specific and list "Jan 1, 2016" ... "Jan 1, 2017"

Done.

> Fig.2: Can you match the minor grid lines with the start of a month, rather than the middle of a month? This will make it much easier to identify the approximate dates for some of the other spikes. No need to label these minor grid lines.

@@

Done.

> Fig.2: You mention "small multiples" only in a figure caption. This concept is a central
part of your calendar graphics and should be summarized in more details (including references) in the main text. Also mention Unwin & Valero-Mora's "Ensemble Graphics", JCGS, 27(1), as a major concept that applies to both sets of related figures in your article.

ToDo

> Fig.3: Add tick marks at 0 and 24. See whether tick mark labels fit. If not, OK to omit those for 0 and 24.

@@

Done.

> p6, l8: You mention ggTimeSeries & ggcal. For completeness, also cite Jones (2016) Calendar Heatmaps, https://rpubs.com/haj3/calheatmap, and possibly Wong's TimeProjection R package, https://cran.r-project.org/web/packages/TimeProjection/index.html

> p6, l49-...: Starting a section called "Data transformation" with an example/figure seems to be strange. Can you first start with the formal steps and then place this example/figure after the formal part, i.e., around p.8, l38.

> Fig.4: Prior to reading the text on p8, I was really confused and even assumed there was a major bug in your R code. Having 2 days in early May and 1 day in early Oct, then 3 days missing, and then the remainder of the month is really confusing for someone who only looks at the figure without reading the text in detail. At least mention this layout anomaly in the figure caption.

Thanks. I have pointed out the layout anomalies for May and October in the Figure 6 caption (previously Figure 4).

> p8, l19: "wrap the last few days up to the top row of the block": This answers
my comment for Fig.4 now, but this layout still remains misleading. We expect to
see some similar temporal pattern in nearby graphs. But there could be considerable differences over a 30-day period, e.g., in your example from Section 3:
What if these 1 or 2 days are after some summer vacation with lots of air conditioning
use, but the vacation already started in the previous month and continued through the middle of the current month. Suddenly, there will be a few huge spikes that interrupt
the overall low-energy pattern. I could think of 2 possible solutions:
(i) Add a 6th week for each month; or (ii) Add the extra days to the
start of the next month. This is sometimes called a
"Calendar Heatmap Tetris Chart", see for example https://stackoverflow.com/questions/27000131/calendar-heat-map-tetris-chart
[both of these features could become additional user options for your R function]

ToDo.

> p8, l44: "Between each month requires some small amount of white space, denoted by b." - strange sentence; rephrase

### 2019-03-05 main.Rmd
revision done for reviewer 3 pg2

In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events. There is some existing work in this area. For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days. It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office. The calendar-based heatmap was implemented in two R packages: **ggTimeSeries** [@R-ggTimeSeries] and **ggcal** [@R-ggcal]. However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day. Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation. Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].

@@

```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016. The disparities between week day and weekend along with public holiday are immediately apparent."}
```

The algorithm for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar. Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block. The notation for creating these cells is as follows:

@@

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$. Between each month requires some small amount of white space, denoted by $b$. Figure \ref{fig:year-diagram} illustrates this layout where $M = 3$ and $N = 4$.

@@

Each cell forms a canvas on which to draw the data. Initialize the canvas to have limits $[0, 1]$ both horizontally and vertically. For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically. Each variable is scaled to have values in $[0,1]$, using the minimum and maximum of all the data values to be displayed, assuming fixed scales. Let $h$ be the scaled hour, and $c$ the scaled count.

@@

>>>>

In contrast, calendar-based graphics unpack the temporal variable, at different resolutions, to digest multiple seasonalities and special events. There is some existing work in this area. For example, @VanWijkCluster1999 developed a calendar view of the heatmap to represent the number of employees in the work place over a year, where colors indicate different clusters derived from the days. It contrasts week days and weekends, highlights public holidays, and presents other known seasonal variation such as school vacations, all of which have influence over the turn-outs in the office. Alongside @Jones-heatmap, @R-TimeProjection, @R-ggTimeSeries, and @R-ggcal implemented some variants of calendar-based heatmaps as in R packages: **TimeProjection**, **ggTimeSeries**, and **ggcal** respectively. However, these techniques are limited to color-encoding graphics and are unable to use time scales smaller than a day. Time of day, which serves as one of the most important aspects in explaining substantial variations arising from the pedestrian sensor data, will be neglected through daily aggregation. Additionally, if simply using colored blocks rather than curves, it may become perceptually difficult to estimate the shape positions and changes, although using curves comes with the cost of more display capacity [@cleveland1984graphical; @lam2007overview].

@@

The algorithm of transforming data for constructing a calendar plot uses linear algebra, similar to that used in the glyph map displays for spatio-temporal data [@Wickham2012glyph]. To make a year long calendar requires cells for days, embedded in blocks corresponding to months, organized into a grid layout for a year. Each month can be captured with 35 (5 $\times$ 7) cells, where the top left is Monday of week 1, and the bottom right is Sunday of week 5 by default. These cells provide a micro canvas on which to plot the data. The first day of the month could be any of Monday--Sunday, which is determined by the year of the calendar. Months are of different lengths, ranging from 28 to 31 days, and each month could extend over six weeks but the convention in these months is to wrap the last few days up to the top row of the block. The notation for creating these cells is as follows:

@@

To create the layout for a full year, $(m, n)$ denotes the position of the month arranged in the plot, where $1 \le m \le M$ and $1 \le n \le N$; $b$ denotes the small amount of white space between each month for visual separation. Figure \ref{fig:year-diagram} illustrates this layout where $M = 3$ and $N = 4$.

@@

Each cell forms a canvas on which to draw the data. Initialize the canvas to have limits $[0, 1]$ both horizontally and vertically. For the pedestrian sensor data, within each cell, hour is plotted horizontally and count is plotted vertically. Each variable is scaled to have values in $[0, 1]$, using the minimum and maximum of all the data values to be displayed, assuming fixed scales. Let $h$ be the scaled hour, and $c$ the scaled count.

@@

```{r fs-2016, fig.height = 7, fig.cap = "The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The disparities between week day and weekend along with public holiday are immediately apparent. The arrangement of the data into a $3 \\times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October."}
```

### 2019-03-05 main.tex
revision done for reviewer 3 pg2

In contrast, calendar-based graphics unpack the temporal variable, at
different resolutions, to digest multiple seasonalities and special
events. There is some existing work in this area. For example,
\citet{VanWijkCluster1999} developed a calendar view of the heatmap to
represent the number of employees in the work place over a year, where
colors indicate different clusters derived from the days. It contrasts
week days and weekends, highlights public holidays, and presents other
known seasonal variation such as school vacations, all of which have
influence over the turn-outs in the office. The calendar-based heatmap
was implemented in two R packages: \textbf{ggTimeSeries}
\citep{R-ggTimeSeries} and \textbf{ggcal} \citep{R-ggcal}. However,
these techniques are limited to color-encoding graphics and are unable
to use time scales smaller than a day. Time of day, which serves as one
of the most important aspects in explaining substantial variations
arising from the pedestrian sensor data, will be neglected through daily
aggregation. Additionally, if simply using colored blocks rather than
curves, it may become perceptually difficult to estimate the shape
positions and changes, although using curves comes with the cost of more
display capacity \citep{cleveland1984graphical, lam2007overview}.

@@

\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. The disparities between week day and weekend along with public holiday are immediately apparent.}\label{fig:fs-2016}
\end{figure}

The algorithm for constructing a calendar plot uses linear algebra,
similar to that used in the glyph map displays for spatio-temporal data
\citep{Wickham2012glyph}. To make a year long calendar requires cells
for days, embedded in blocks corresponding to months, organized into a
grid layout for a year. Each month can be captured with 35 (5 \(\times\)
7) cells, where the top left is Monday of week 1, and the bottom right
is Sunday of week 5 by default. These cells provide a micro canvas on
which to plot the data. The first day of the month could be any of
Monday--Sunday, which is determined by the year of the calendar. Months
are of different lengths, ranging from 28 to 31 days, and each month
could extend over six weeks but the convention in these months is to
wrap the last few days up to the top row of the block. The notation for
creating these cells is as follows:

@@

To create the layout for a full year, \((m, n)\) denotes the position of
the month arranged in the plot, where \(1 \le m \le M\) and
\(1 \le n \le N\). Between each month requires some small amount of
white space, denoted by \(b\). Figure \ref{fig:year-diagram} illustrates
this layout where \(M = 3\) and \(N = 4\).

@@

Each cell forms a canvas on which to draw the data. Initialize the
canvas to have limits \([0, 1]\) both horizontally and vertically. For
the pedestrian sensor data, within each cell, hour is plotted
horizontally and count is plotted vertically. Each variable is scaled to
have values in \([0,1]\), using the minimum and maximum of all the data
values to be displayed, assuming fixed scales. Let \(h\) be the scaled
hour, and \(c\) the scaled count.

@@

>>>>

In contrast, calendar-based graphics unpack the temporal variable, at
different resolutions, to digest multiple seasonalities and special
events. There is some existing work in this area. For example,
\citet{VanWijkCluster1999} developed a calendar view of the heatmap to
represent the number of employees in the work place over a year, where
colors indicate different clusters derived from the days. It contrasts
week days and weekends, highlights public holidays, and presents other
known seasonal variation such as school vacations, all of which have
influence over the turn-outs in the office. Alongside
\citet{Jones-heatmap}, \citet{R-TimeProjection}, \citet{R-ggTimeSeries},
and \citet{R-ggcal} implemented some variants of calendar-based heatmaps
as in R packages: \textbf{TimeProjection}, \textbf{ggTimeSeries}, and
\textbf{ggcal} respectively. However, these techniques are limited to
color-encoding graphics and are unable to use time scales smaller than a
day. Time of day, which serves as one of the most important aspects in
explaining substantial variations arising from the pedestrian sensor
data, will be neglected through daily aggregation. Additionally, if
simply using colored blocks rather than curves, it may become
perceptually difficult to estimate the shape positions and changes,
although using curves comes with the cost of more display capacity
\citep{cleveland1984graphical, lam2007overview}.

@@

The algorithm of transforming data for constructing a calendar plot uses
linear algebra, similar to that used in the glyph map displays for
spatio-temporal data \citep{Wickham2012glyph}. To make a year long
calendar requires cells for days, embedded in blocks corresponding to
months, organized into a grid layout for a year. Each month can be
captured with 35 (5 \(\times\) 7) cells, where the top left is Monday of
week 1, and the bottom right is Sunday of week 5 by default. These cells
provide a micro canvas on which to plot the data. The first day of the
month could be any of Monday--Sunday, which is determined by the year of
the calendar. Months are of different lengths, ranging from 28 to 31
days, and each month could extend over six weeks but the convention in
these months is to wrap the last few days up to the top row of the
block. The notation for creating these cells is as follows:

@@

To create the layout for a full year, \((m, n)\) denotes the position of
the month arranged in the plot, where \(1 \le m \le M\) and
\(1 \le n \le N\); \(b\) denotes the small amount of white space between
each month for visual separation. Figure \ref{fig:year-diagram}
illustrates this layout where \(M = 3\) and \(N = 4\).

@@

Each cell forms a canvas on which to draw the data. Initialize the
canvas to have limits \([0, 1]\) both horizontally and vertically. For
the pedestrian sensor data, within each cell, hour is plotted
horizontally and count is plotted vertically. Each variable is scaled to
have values in \([0, 1]\), using the minimum and maximum of all the data
values to be displayed, assuming fixed scales. Let \(h\) be the scaled
hour, and \(c\) the scaled count.

@@

\caption{The calendar-based display of hourly foot traffic at Flagstaff Station using line glyphs. The disparities between week day and weekend along with public holiday are immediately apparent. The arrangement of the data into a $3 \times 4$ monthly grid represents all the traffic in 2016. Note that the algorithm wraps the last few days in the sixth week to the top row of each month block for a compact layout, which occurs to May and October.}\label{fig:fs-2016}
\end{figure}
